---
title: "Coinoculation"
author: "James C. Kosmopoulos"
date: "`r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r Load packages, message=FALSE, warning=FALSE}
# global options
knitr::opts_chunk$set(echo = TRUE)

# set working directory
setwd("~/Documents/Research/co-inoculation/coinoculation")

# load packages
library("tidyverse") # includes ggplot2, dplyr, readr, stringr
library("knitr") # produce the knitted doc
library("emmeans") # for posthoc data analysis
library("lme4") # linear mixed models
library("car") # ANOVA function for linear mixed models
library("gridExtra") # for arranging plots
library("cowplot")
```


# Summarize,organize, and save cleaned data

```{r, echo = FALSE, warning = FALSE}
tp_raw <- read_csv("all_data.csv")

# filter function to remove outliers
tp_all <- tp_raw %>% filter(!sample %in% c("InCtrl-2-r15", "733-17-r27", "717-6-r33")) %>% droplevels()
# create shoot:root ratio
tp_all$ratio <-tp_all$shoot_mass / tp_all$root_mass
#re-arrange order of treatments to make some figure look nicer
tp_all$treatment <- factor(tp_all$treatment, level = c("522", "702", "717", "733", "141", "522co141", "702co141", "717co141", "733co141", "InCtrl", "ExCtrl"))
# add in type col (to compare internal, external, and treated)
tp_all$type <- ifelse(tp_all$treatment == "InCtrl", "internal",
                    ifelse(tp_all$treatment == "ExCtrl", "external","treated"))
tp_all$type <- factor(tp_all$type, levels = c("internal","external","treated"))

#create column designating treatment group
tp_all$group <- ifelse(tp_all$treatment == "141", "Ensifer-only",
                       ifelse(tp_all$treatment == "522" , "single",
                              ifelse(tp_all$treatment == "702", "single",
                                     ifelse(tp_all$treatment == "717", "single",
                                            ifelse(tp_all$treatment == "733", "single",
                                                   ifelse(tp_all$treatment == "InCtrl", "control",
                                                          ifelse(tp_all$treatment == "ExCtrl", "control", "co-inoculated")))))))


tp_all_sum <- tp_all %>%
  group_by(type) %>%
  summarize(count = n(),
            # timepoint1 data
            # height
            mean_height1 = mean(height1, na.rm = TRUE), sd_height1 = sd(height1, na.rm = TRUE),
            SE_height1 = sd_height1/sqrt(count),
            # chlorophyll
            mean_chloro1 = mean(chloro1, na.rm = TRUE), sd_chloro1 = sd(chloro1, na.rm = TRUE),
            SE_chloro1 = sd_chloro1/sqrt(count),
            # leaf
            mean_leaf1 = mean(leaf1, na.rm = TRUE), sd_leaf1 = sd(leaf1, na.rm = TRUE),
            SE_leaf1 = sd_leaf1/sqrt(count),

            # timepoint2 data
            # height
            mean_height2 = mean(height2, na.rm = TRUE), sd_height2 = sd(height2, na.rm = TRUE),
            SE_height2 = sd_height2/sqrt(count),
            # chlorophyll
            mean_chloro2 = mean(chloro2, na.rm = TRUE), sd_chloro2 = sd(chloro2, na.rm = TRUE),
            SE_chloro2 = sd_chloro2/sqrt(count),
            # leaf
            mean_leaf2 = mean(leaf2, na.rm = TRUE), sd_leaf2 = sd(leaf2, na.rm = TRUE),
            SE_leaf2 = sd_leaf2/sqrt(count),

            # timepoint3 data
            # height
            mean_height3 = mean(height3, na.rm = TRUE), sd_height3 = sd(height3, na.rm = TRUE),
            SE_height3 = sd_height3/sqrt(count),
            # chlorophyll
            mean_chloro3 = mean(chloro3, na.rm = TRUE), sd_chloro3 = sd(chloro3, na.rm = TRUE),
            SE_chloro3 = sd_chloro3/sqrt(count),
            # leaf
            mean_leaf3 = mean(leaf3, na.rm = TRUE), sd_leaf3 = sd(leaf3, na.rm = TRUE),
            SE_leaf3 = sd_leaf3/sqrt(count),
            
            # timepoint4 data
            # height
            mean_height4 = mean(height4, na.rm = TRUE), sd_height4 = sd(height4, na.rm = TRUE),
            SE_height4 = sd_height4/sqrt(count),
            # chlorophyll
            mean_chloro4 = mean(chloro4, na.rm = TRUE), sd_chloro4 = sd(chloro4, na.rm = TRUE),
            SE_chloro4 = sd_chloro4/sqrt(count),
            # leaf
            mean_leaf4 = mean(leaf4, na.rm = TRUE), sd_leaf4 = sd(leaf4, na.rm = TRUE),
            SE_leaf4 = sd_leaf4/sqrt(count),
            
            # nodule dissection data
            # mean nodule mass per tube
            mean_nod_mass = mean(nod_mass, na.rm = TRUE),
            sd_nod_mass = sd(nod_mass, na.rm = TRUE),
            SE_nod_mass = sd_nod_mass/sqrt(count),
            # total nodules per tube
            mean_total_nod = mean(total_nod, na.rm = TRUE),
            sd_total_nod = sd(total_nod, na.rm = TRUE),
            SE_total_nod = sd_total_nod/sqrt(count),
            
            # biomass
            # mean shoot biomass (mg)
            mean_shoot_mass = mean(shoot_mass, na.rm = TRUE),
            sd_shoot_mass = sd(shoot_mass, na.rm = TRUE),
            SE_shoot_mass = sd_shoot_mass/sqrt(count),
            # mean root biomass (mg)
            mean_root_mass = mean(root_mass, na.rm = TRUE),
            sd_root_mass = sd(root_mass, na.rm = TRUE),
            SE_root_mass = sd_root_mass/sqrt(count),
            # ratio
            mean_ratio = mean(ratio, na.rm = TRUE),
            sd_ratio = sd(ratio, na.rm = TRUE),
            SE_ratio = sd_ratio/sqrt(count))
            # when adding more data, make sure to delete one of the ')' above in (count)) and add a comma

tp_all_long <- gather(tp_all, key = "trait", value = "measurement", height1:ratio)# use gather to stack the trait columns

tp_long_sum <- tp_all_long %>%
  group_by(type, trait) %>%
  summarize(count = n(),
            mean_trait = mean(measurement, na.rm = TRUE), sd_trait = sd(measurement, na.rm = TRUE),
            SE_trait = sd_trait/sqrt(count))
tp_long_sum$trait <- factor(tp_long_sum$trait, level = c("height1", "height2", "height3", "height4", "leaf1", "leaf2", "leaf3", "leaf4", "chloro1", "chloro2", "chloro3", "chloro4", "shoot_mass", "root_mass", "ratio", "nod_mass", "total_nod"))

# this includes controls
tp_long_sum_controlsplus <- tp_all_long %>%
  group_by(treatment, trait) %>%
  summarize(count = n(),
            mean_trait = mean(measurement, na.rm = TRUE), sd_trait = sd(measurement, na.rm = TRUE),
            SE_trait = sd_trait/sqrt(count))

tp_treated <- tp_all %>% # for ANOVA
  filter(type == "treated") %>%
  droplevels(.)

tp_treated_long <- gather(tp_treated, key = "trait", value = "measurement", height1:ratio)
tp_treated_long$trait <- factor(tp_treated_long$trait, level = c("height1", "height2", "height3", "height4", "leaf1", "leaf2", "leaf3", "leaf4", "chloro1", "chloro2", "chloro3", "chloro4", "shoot_mass", "root_mass", "ratio", "nod_mass", "total_nod"))

tp_treated_long_sum <- tp_treated_long %>%
  group_by(treatment, trait) %>%
  summarize(count = n(),
            mean_trait = mean(measurement, na.rm = TRUE), sd_trait = sd(measurement, na.rm = TRUE),
            SE_trait = sd_trait/sqrt(count))

#groups for chlorophyll through time
tp_treated_long_sum$chloro.time <- ifelse(tp_treated_long_sum$treatment == "141", "A", "B")
tp_treated_long_sum$timepoint <- ifelse(tp_treated_long_sum$trait == "height1", "1",
                    ifelse(tp_treated_long_sum$trait == "leaf1", "1",
                      ifelse(tp_treated_long_sum$trait == "chloro1", "1",
                        ifelse(tp_treated_long_sum$trait == "height2", "2",
                          ifelse(tp_treated_long_sum$trait == "leaf2", "2",
                            ifelse(tp_treated_long_sum$trait == "chloro2", "2",
                              ifelse(tp_treated_long_sum$trait == "height3", "3",
                                ifelse(tp_treated_long_sum$trait == "leaf3", "3",
                                  ifelse(tp_treated_long_sum$trait == "chloro3", "3", "4")))))))))

#groups for chlorophyll through time
tp_long_sum_controlsplus$chloro.time <- ifelse(tp_long_sum_controlsplus$treatment == "141", "A", "B")
tp_long_sum_controlsplus$timepoint <- ifelse(tp_long_sum_controlsplus$trait == "height1", "1",
                    ifelse(tp_long_sum_controlsplus$trait == "leaf1", "1",
                      ifelse(tp_long_sum_controlsplus$trait == "chloro1", "1",
                        ifelse(tp_long_sum_controlsplus$trait == "height2", "2",
                          ifelse(tp_long_sum_controlsplus$trait == "leaf2", "2",
                            ifelse(tp_long_sum_controlsplus$trait == "chloro2", "2",
                              ifelse(tp_long_sum_controlsplus$trait == "height3", "3",
                                ifelse(tp_long_sum_controlsplus$trait == "leaf3", "3",
                                  ifelse(tp_long_sum_controlsplus$trait == "chloro3", "3", "4")))))))))
```

# Plot: chloro through time
```{r}
# create means for comparison
table.nonensifer <- tp_treated_long_sum %>% filter(lmm1 %in% c("non-Ensifer"), na.rm = TRUE)
table.nonensifer.chloro1 <- mean.nonensifer %>% filter(trait %in% c("chloro1"), na.rm = TRUE)
mean.nonensifer.chloro1 <- mean(table.nonensifer.chloro1$mean_trait)
mean.nonensifer.chloro1
table.nonensifer <- tp_treated_long_sum %>% filter(lmm1 %in% c("non-Ensifer"), na.rm = TRUE)
table.nonensifer.chloro2 <- mean.nonensifer %>% filter(trait %in% c("chloro2"), na.rm = TRUE)
mean.nonensifer.chloro2 <- mean(table.nonensifer.chloro2$mean_trait)
mean.nonensifer.chloro2
table.nonensifer <- tp_treated_long_sum %>% filter(lmm1 %in% c("non-Ensifer"), na.rm = TRUE)
table.nonensifer.chloro3 <- mean.nonensifer %>% filter(trait %in% c("chloro3"), na.rm = TRUE)
mean.nonensifer.chloro3 <- mean(table.nonensifer.chloro3$mean_trait)
mean.nonensifer.chloro3
table.nonensifer <- tp_treated_long_sum %>% filter(lmm1 %in% c("non-Ensifer"), na.rm = TRUE)
table.nonensifer.chloro4 <- mean.nonensifer %>% filter(trait %in% c("chloro4"), na.rm = TRUE)
mean.nonensifer.chloro4 <- mean(table.nonensifer.chloro4$mean_trait)
mean.nonensifer.chloro4
table.141 <- tp_treated_long_sum %>% filter(lmm1 %in% c("141"), na.rm = TRUE)
table.141.chloro1 <- table.141 %>% filter(trait %in% c("chloro1"), na.rm = TRUE)
mean.141.chloro1 <- mean(table.141.chloro1$mean_trait)
mean.141.chloro1
table.141 <- tp_treated_long_sum %>% filter(lmm1 %in% c("141"), na.rm = TRUE)
table.141.chloro2 <- table.141 %>% filter(trait %in% c("chloro2"), na.rm = TRUE)
mean.141.chloro2 <- mean(table.141.chloro2$mean_trait)
mean.141.chloro2
table.141 <- tp_treated_long_sum %>% filter(lmm1 %in% c("141"), na.rm = TRUE)
table.141.chloro3 <- table.141 %>% filter(trait %in% c("chloro3"), na.rm = TRUE)
mean.141.chloro3 <- mean(table.141.chloro3$mean_trait)
mean.141.chloro3
table.141 <- tp_treated_long_sum %>% filter(lmm1 %in% c("141"), na.rm = TRUE)
table.141.chloro4 <- table.141 %>% filter(trait %in% c("chloro4"), na.rm = TRUE)
mean.141.chloro4 <- mean(table.141.chloro4$mean_trait)
mean.141.chloro4
table.coinoculation <- tp_treated_long_sum %>% filter(lmm1 %in% c("co-inoculation"), na.rm = TRUE)
table.coinoculation.chloro1 <- table.coinoculation %>% filter(trait %in% c("chloro1"), na.rm = TRUE)
mean.coinoculation.chloro1 <- mean(table.coinoculation.chloro1$mean_trait)
mean.coinoculation.chloro1
table.coinoculation <- tp_treated_long_sum %>% filter(lmm1 %in% c("co-inoculation"), na.rm = TRUE)
table.coinoculation.chloro2 <- table.coinoculation %>% filter(trait %in% c("chloro2"), na.rm = TRUE)
mean.coinoculation.chloro2 <- mean(table.coinoculation.chloro2$mean_trait)
mean.coinoculation.chloro2
table.coinoculation <- tp_treated_long_sum %>% filter(lmm1 %in% c("co-inoculation"), na.rm = TRUE)
table.coinoculation.chloro3 <- table.coinoculation %>% filter(trait %in% c("chloro3"), na.rm = TRUE)
mean.coinoculation.chloro3 <- mean(table.coinoculation.chloro3$mean_trait)
mean.coinoculation.chloro3
table.coinoculation <- tp_treated_long_sum %>% filter(lmm1 %in% c("co-inoculation"), na.rm = TRUE)
table.coinoculation.chloro4 <- table.coinoculation %>% filter(trait %in% c("chloro4"), na.rm = TRUE)
mean.coinoculation.chloro4 <- mean(table.coinoculation.chloro4$mean_trait)
mean.coinoculation.chloro4
chloro.means = c(mean.nonensifer.chloro1, mean.nonensifer.chloro2, mean.nonensifer.chloro3, mean.nonensifer.chloro4, mean.141.chloro1, mean.141.chloro2, mean.141.chloro3, mean.141.chloro4, mean.coinoculation.chloro1, mean.coinoculation.chloro2, mean.coinoculation.chloro3, mean.coinoculation.chloro4)
chloro.time.2.table <- as.data.frame(x = chloro.means)
chloro.means.lmm1 <- c("non-Ensifer", "non-Ensifer", "non-Ensifer", "non-Ensifer", "Ensifer", "Ensifer", "Ensifer", "Ensifer", "co-inoculation", "co-inoculation", "co-inoculation", "co-inoculation")
chloro.time.2.table$group <- chloro.means.lmm1
chloro.timepoints <- c(1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4)
chloro.time.2.table$timepoint <- chloro.timepoints
chloro.time.2.table$group <- factor(chloro.time.2.table$group, levels = c("non-Ensifer", "co-inoculation", "Ensifer"))

# Now plot it
fig.chloro.thru.time.2 <- ggplot(chloro.time.2.table, aes(x=timepoint, y=chloro.means)) + 
  geom_point(size = 5, aes(color = group)) +
  geom_line(aes(group = group, color = group)) +
  scale_color_manual(values = c("#CC79A7", "#D55E00", "#0072B2", "#F0E442", "#009E73", "#56B4E9", "#E69F00", "#000000")) +
  scale_shape_manual(values = c(4, 4, 4, 4, 4, 4)) +
  annotate("text", x = 1, y = 39.3, 
              label = "c",
              fontface = 1, size = 5) +
  annotate("text", x = 1, y = 36, 
              label = "ab",
              fontface = 1, size = 5) +
  annotate("text", x = 1, y = 31.65, 
              label = "a",
              fontface = 1, size = 5) +
  xlab("Weeks after inoculation") +
  ylab("Chlorophyll content (SPAD)") +
  ggtitle("Average chlorophyll content through time") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "right")
plot(fig.chloro.thru.time.2)
save_plot("./figures/chloro_thru_time2.png", fig.chloro.thru.time.2, base_asp = 2.25)
```
# Control t-tests
```{r}
tp_control <- tp_all %>% filter(treatment %in% c("InCtrl", "ExCtrl")) %>% droplevels()
t.test(tp_control$shoot_mass ~ tp_control$treatment) # p=0.6460 InCtrl and ExCtrl not sig diff.
t.test(tp_control$root_mass ~ tp_control$treatment) # p=0.4964 InCtrl and ExCtrl not sig diff.
t.test(tp_control$chloro4 ~ tp_control$treatment) # p= 06061 InCtrl and ExCtrl not sig diff.
t.test(tp_control$leaf4 ~ tp_control$treatment) # p=0.3871  InCtrl and ExCtrl not sig diff.
t.test(tp_control$height4 ~ tp_control$treatment) # p=0.1903 InCtrl and ExCtrl not sig diff.
```
# Lmm with controls to non-ensifer (direct effects)... tests direct effects of non-Ensifer on traits compared to controls
```{r}
tp_direct <- tp_all %>% filter(treatment %in% c("InCtrl", "ExCtrl", "522", "702", "717", "733")) %>% droplevels()
tp_direct$lmm <- ifelse(tp_direct$treatment == "InCtrl" | 
                        tp_direct$treatment == "ExCtrl", "control", "non-Ensifer")

lmm.ctrl.shoot_mass <- lmer(shoot_mass ~  lmm + (1|rack), data = tp_direct)
summary(lmm.ctrl.shoot_mass)
(anova_lmm.ctrl.shoot_mass <- Anova(lmm.ctrl.shoot_mass, type =2)) # p=0.6913

lmm.ctrl.root_mass <- lmer(root_mass ~  lmm + (1|rack), data = tp_direct)
summary(lmm.ctrl.root_mass)
(anova_lmm.ctrl.root_mass <- Anova(lmm.ctrl.root_mass, type =2)) # p=0.9800

#don't include total nod or nod mass because none of the plants in these data had nodules
#lmm.ctrl.total_nod <- lmer(total_nod ~  lmm + (1|rack), data = tp_direct)
#summary(lmm.ctrl.total_nod)
#(anova_lmm.ctrl.total_nod <- Anova(lmm.ctrl.total_nod, type =2))

#lmm.ctrl.nod_mass <- lmer(nod_mass ~  lmm + (1|rack), data = tp_direct)
#summary(lmm.ctrl.nod_mass)
#(anova_lmm.ctrl.nod_mass <- Anova(lmm.ctrl.nod_mass, type =2))

lmm.ctrl.height4 <- lmer(height4 ~  lmm + (1|rack), data = tp_direct)
summary(lmm.ctrl.height4)
(anova_lmm.ctrl.height4 <- Anova(lmm.ctrl.height4, type =2)) # p=0.4519

lmm.ctrl.chloro4 <- lmer(chloro4 ~  lmm + (1|rack), data = tp_direct)
summary(lmm.ctrl.chloro4)
(anova_lmm.ctrl.chloro4 <- Anova(lmm.ctrl.chloro4, type =2)) # p=0.8651

lmm.ctrl.leaf4 <- lmer(leaf4 ~  lmm + (1|rack), data = tp_direct)
summary(lmm.ctrl.leaf4)
(anova_lmm.ctrl.leaf4 <- Anova(lmm.ctrl.leaf4, type =2)) # p=0.9345
```
# Lmm with Ensifer vs co-inoculated, fdr djustment
```{r}
tp_non.ens.rm <- tp_all %>% filter(treatment %in% c("141", "522co141", "702co141", "717co141", "733co141")) %>% droplevels()
tp_non.ens.rm$lmm <- ifelse(tp_non.ens.rm$treatment == "522co141" |
                              tp_non.ens.rm$treatment == "702co141" |
                              tp_non.ens.rm$treatment == "717co141" |
                              tp_non.ens.rm$treatment == "733co141", "co-inoculated", "Ensifer")

lmm.non.ens.rm.shoot_mass <- lmer(shoot_mass ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.shoot_mass)
(anova_lmm.non.ens.rm.shoot_mass <- Anova(lmm.non.ens.rm.shoot_mass, type = 2))
lmm.non.ens.rm.shoot_mass.emms <- emmeans(lmm.non.ens.rm.shoot_mass, pairwise ~ lmm)
lmm.non.ens.rm.shoot_mass.emms.sum <- summary(lmm.non.ens.rm.shoot_mass.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=0.101, not sig

lmm.non.ens.rm.total_nod <- lmer(total_nod ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.total_nod)
(anova_lmm.non.ens.rm.total_nod <- Anova(lmm.non.ens.rm.total_nod, type = 2))
lmm.non.ens.rm.total_nod.emms <- emmeans(lmm.non.ens.rm.total_nod, pairwise ~ lmm)
lmm.non.ens.rm.total_nod.emms.sum <- summary(lmm.non.ens.rm.total_nod.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=.047, sig!
plot(lmm.non.ens.rm.total_nod)

lmm.non.ens.rm.root_mass <- lmer(root_mass ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.root_mass)
(anova_lmm.non.ens.rm.root_mass <- Anova(lmm.non.ens.rm.root_mass, type = 2))
lmm.non.ens.rm.root_mass.emms <- emmeans(lmm.non.ens.rm.root_mass, pairwise ~ lmm)
lmm.non.ens.rm.root_mass.emms.sum <- summary(lmm.non.ens.rm.root_mass.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=0.187 not sig

lmm.non.ens.rm.chloro4 <- lmer(chloro4 ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.chloro4)
(anova_lmm.non.ens.rm.chloro4 <- Anova(lmm.non.ens.rm.chloro4, type = 2))
lmm.non.ens.rm.chloro4.emms <- emmeans(lmm.non.ens.rm.chloro4, pairwise ~ lmm)
lmm.non.ens.rm.chloro4.emms.sum <- summary(lmm.non.ens.rm.chloro4.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=0.239 not sig

lmm.non.ens.rm.leaf4 <- lmer(leaf4 ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.leaf4)
(anova_lmm.non.ens.rm.leaf4 <- Anova(lmm.non.ens.rm.leaf4, type = 2))
lmm.non.ens.rm.leaf4.emms <- emmeans(lmm.non.ens.rm.leaf4, pairwise ~ lmm)
lmm.non.ens.rm.leaf4.emms.sum <- summary(lmm.non.ens.rm.leaf4.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=0.110 not sig

lmm.non.ens.rm.height4 <- lmer(height4 ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.height4)
(anova_lmm.non.ens.rm.height4 <- Anova(lmm.non.ens.rm.height4, type = 2))
lmm.non.ens.rm.height4.emms <- emmeans(lmm.non.ens.rm.height4, pairwise ~ lmm)
lmm.non.ens.rm.height4.emms.sum <- summary(lmm.non.ens.rm.height4.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=0.559 not sig

lmm.non.ens.rm.nod_mass <- lmer(nod_mass ~ lmm + (1|rack), data = tp_non.ens.rm)
summary(lmm.non.ens.rm.nod_mass)
(anova_lmm.non.ens.rm.nod_mass <- Anova(lmm.non.ens.rm.nod_mass, type = 2))
lmm.non.ens.rm.nod_mass.emms <- emmeans(lmm.non.ens.rm.nod_mass, pairwise ~ lmm)
lmm.non.ens.rm.nod_mass.emms.sum <- summary(lmm.non.ens.rm.nod_mass.emms, infer = c(TRUE,TRUE), adjust = "fdr") # p=0.118 not sig
```

# Lmm with all data, pairwise contrasts for every treatment/control, fdr djustment
```{r}
lmm.all.shoot_mass <- lmer(shoot_mass ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.shoot_mass)
(anova_lmm.all.shoot_mass <- Anova(lmm.all.shoot_mass, type = 2))
lmm.all.shoot_mass.emms <- emmeans(lmm.all.shoot_mass, pairwise ~ treatment)
lmm.all.shoot_mass.emms.sum <- summary(lmm.all.shoot_mass.emms, infer = c(TRUE,TRUE), adjust = "fdr") # 717-141 and 733-141 sig!
plot(lmm.all.shoot_mass)

lmm.all.total_nod <- lmer(total_nod ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.total_nod)
(anova_lmm.all.total_nod <- Anova(lmm.all.total_nod, type = 2))
lmm.all.total_nod.emms <- emmeans(lmm.all.total_nod, pairwise ~ treatment)
lmm.all.total_nod.emms.sum <- summary(lmm.all.total_nod.emms, infer = c(TRUE,TRUE), adjust = "fdr") # 717-141 and 733-141 sig!
plot(lmm.all.total_nod)

lmm.all.root_mass <- lmer(root_mass ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.root_mass)
(anova_lmm.all.root_mass <- Anova(lmm.all.root_mass, type = 2))
lmm.all.root_mass.emms <- emmeans(lmm.all.root_mass, pairwise ~ treatment)
lmm.all.root_mass.emms.sum <- summary(lmm.all.root_mass.emms, infer = c(TRUE,TRUE), adjust = "fdr") # not sig
plot(lmm.all.root_mass)

lmm.all.chloro4 <- lmer(chloro4 ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.chloro4)
(anova_lmm.all.chloro4 <- Anova(lmm.all.chloro4, type = 2))
lmm.all.chloro4.emms <- emmeans(lmm.all.chloro4, pairwise ~ treatment)
lmm.all.chloro4.emms.sum <- summary(lmm.all.chloro4.emms, infer = c(TRUE,TRUE), adjust = "fdr") # not sig
plot(lmm.all.chloro4)

lmm.all.leaf4 <- lmer(leaf4 ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.leaf4)
(anova_lmm.all.leaf4 <- Anova(lmm.all.leaf4, type = 2))
lmm.all.leaf4.emms <- emmeans(lmm.all.leaf4, pairwise ~ treatment)
lmm.all.leaf4.emms.sum <- summary(lmm.all.leaf4.emms, infer = c(TRUE,TRUE), adjust = "fdr") # 717-141 and 733-141 sig!
plot(lmm.all.leaf4)

lmm.all.height4 <- lmer(height4 ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.height4)
(anova_lmm.all.height4 <- Anova(lmm.all.height4, type = 2))
lmm.all.height4.emms <- emmeans(lmm.all.height4, pairwise ~ treatment)
lmm.all.height4.emms.sum <- summary(lmm.all.height4.emms, infer = c(TRUE,TRUE), adjust = "fdr") # not sig
plot(lmm.all.height4)

lmm.all.nod_mass <- lmer(nod_mass ~ treatment + (1|rack), data = tp_all)
summary(lmm.all.nod_mass)
(anova_lmm.all.nod_mass <- Anova(lmm.all.nod_mass, type = 2))
lmm.all.nod_mass.emms <- emmeans(lmm.all.nod_mass, pairwise ~ treatment)
lmm.all.nod_mass.emms.sum <- summary(lmm.all.nod_mass.emms, infer = c(TRUE,TRUE), adjust = "fdr") # not sig
plot(lmm.all.nod_mass)
```

# P-value table for lmm with all data
```{r}

```


# Plot: shoot mass
```{r}
# All treatments and controls
(boxplot.shoot_mass.all <- ggplot(tp_all , aes(x=treatment, y=shoot_mass)) +
  geom_boxplot() +
  xlab("Treatment") +
  ylab("Shoot mass (g)") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.shoot_mass.all)
save_plot("./figures/boxplot_shoot_mass_all.png", boxplot.shoot_mass.all, base_asp = 3)

# 141 and co-inoculated only
(boxplot.shoot_mass <- ggplot(tp_all %>% filter(treatment == c("141", "522co141", "702co141", "717co141", "733co141")), aes(x=treatment, y=shoot_mass)) +
  geom_boxplot() +
  xlab("Treatment") +
  ylab("Shoot mass (g)") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.shoot_mass.all)
save_plot("./figures/boxplot_shoot_mass.png", boxplot.shoot_mass, base_asp = 2)
```

# Plot: leaf number
```{r}
# All treatments and controls
(boxplot.leaf.all <- ggplot(tp_all, aes(x=treatment, y=leaf4)) +
  geom_boxplot() +
  xlab("Treatment") +
  ylab("Leaf count") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.leaf.all)
save_plot("./figures/boxplot_leaf_all.png", boxplot.leaf.all, base_asp = 3)

# 141 and co-inoculated only
(boxplot.leaf <- ggplot(tp_all %>% filter(treatment == c("141", "522co141", "702co141", "717co141", "733co141")), aes(x=treatment, y=leaf4)) +
  geom_boxplot() +
  xlab("Treatment") +
  ylab("Leaf count") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.leaf)
save_plot("./figures/boxplot_leaf.png", boxplot.leaf, base_asp = 2)
```


# Plot: nodule count
```{r}
# boxplot, 141 and co-inoculated treatments only
(boxplot.total_nod <- ggplot(tp_all %>% filter(treatment == c("141", "522co141", "702co141", "717co141", "733co141")), aes(x=treatment, y=total_nod)) +
  geom_boxplot() +
  xlab("Treatment") +
  ylab("Nodule count") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.total_nod.nErm)
save_plot("./figures/boxplot_nodule_count.png", boxplot.total_nod, base_asp = 2)

# boxplot, all treatments and controls
(boxplot.total_nod.all <- ggplot(tp_all, aes(x=treatment, y=total_nod)) +
  geom_boxplot() +
  xlab("Treatment") +
  ylab("Nodule count") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.total_nod.all)
save_plot("./figures/boxplot_nodule_count_all.png", boxplot.total_nod.all, base_asp = 3)

#boxplot, treatment group
(boxplot.total_nod.group <- ggplot(tp_all %>% filter(group == c("single", "co-inoculated", "Ensifer-only")), aes(x=group, y=total_nod)) +
  geom_boxplot() +
  xlab("Inoculation group") +
  ylab("Nodule count") +
  #ggtitle("Average nodule count per plant at week 4") +
  theme_minimal_hgrid() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
  )
plot(boxplot.total_nod.group)
save_plot("./figures/boxplot_nodule_count_group.png", boxplot.total_nod.group, base_asp = 2)

```
